---
title: "KLING et al P&N"
author: "Katherine Kling & Carter W. Daniels"
date: "2023-05-11"
output: word_document
---

```{r setup, include=FALSE}
librarian::shelf(DHARMa,
                 dplyr,
                 geiger,
                 ggeffects,
                 ggforestplot,
                 ggplot2,
                 gtools,
                 glmmTMB,
                 Hmisc,
                 interactions,
                 jtools,
                 lme4,
                 marginaleffects,
                 mfx,
                 MuMIn,
                 performance,
                 readr,
                 sjPlot,
                 sys,
                 vcd,
                 vcdExtra) # use to install & load packages, as needed

```

```{r, include = F}
# Necessary Functions

# [1] For model selection of analyses of population & distance
fun.mod.notime.aicc <- function(x) {
  ## TWO PREDICTORS
  # Pop&Dist Int
  x.1<-  AICc(glmmTMB(df2_notime[,x] ~ pop*dist + (1|site:plot.id), family = poisson, data = df2_notime))
  # Pop&Dist
  x.2 <-  AICc(glmmTMB(df2_notime[,x] ~ pop + dist + (1|site:plot.id), family = poisson, data = df2_notime))
  ## SINGLE PREDICTORS
  # Pop Only
  x.3 <-  AICc(glmmTMB(df2_notime[,x] ~ pop + (1|site:plot.id), family = poisson, data = df2_notime))
  # Dist Only
  x.4 <-  AICc(glmmTMB(df2_notime[,x] ~ dist + (1|site:plot.id), family = poisson, data = df2_notime))
  ## NULL MODEL
  # Null
  x.5 <- AICc(glmmTMB(df2_notime[,x] ~ 1 + (1|site:plot.id), family = poisson, data = df2_notime))
  aicc <- c(x.1, x.2, x.3, x.4, x.5)
  names(aicc) <- c("Full-Int", "Full-NoInt", "Pop", "Dist", "Null")
  aicc.w <<- aicw(aicc)
} # Where x is the column # for the dependent variable (e.g., 10 = wood.ct)

## Formulas for assessing the likelihood of each model as compared to the null
# [2] POPULATION & DISTANCE ANALYSES: To calculate likelihood of selected best model compared to null. Returns (1) dAICc (for AICc change between best & null model) & (2) likelihood (exp(dAICc/2))
fun.mod.e2.null <- function(x) {
  dAICc <- aicc.w$delta[x]
  likelihood <- exp(dAICc/2)
  x.1 <- data.frame(dAICc, likelihood)
  x.1
} # where x is the row number for null model values, typically 5 

# [3] Generates null model for change over time analyses
fun.m0t <- function(x) {
  m0 <- glmmTMB(df3_time[,x] ~ 1 + (1|site:plot.id), family = poisson, data = df3_time)
  m0
} # Where x is the column number for the metric of interest in df3_time

# [4] CHANGE OVER TIME ANALYSES: To calculate likelihood & weight of selected best model compared to null. 
fun.mod.e2.null.t <- function (x, y) {
  AICc.null <- AICc(fun.m0t(y))[1]
  AICc.mod <- AICc(x)[1]
  AICc.list <- c(AICc.mod, AICc.null)
  dAICc <- AICc.null - AICc.mod
  likelihood <- exp(dAICc/2)
  weight <- aicw(AICc.list)[1,3]
  x.1 <- data.frame(dAICc, likelihood, weight)
  x.1
} # where x is name of change over time model (e.g., mod.time.timc) & x is the column number for the metric of interest in df3_time (uses function above - fun.m0t)

## Shortcuts for Visualization
# [5] Generate model-predicted values & save in data frame
fun.mod.predict <- function(x) {
  df.predict$predicted <- exp(predict(x, newdata = df.predict, allow.new.levels = T))
  df.predict$ave_predicted <- ave(df.predict$predicted, df.predict$dist, df.predict$pop, FUN=mean)
  
  x.1 <- unique(df.predict[,c("dist","pop","ave_predicted")])
  colnames(x.1)<-c("new_dist","new_pop","ave_predicted")
  x.2 <- x.1$ave_predicted
  x.2
} # Where x = model

# [6] Plot model-predicted lines by population, colored by distance
fun.plot.mod <- function(x,y) {
  y.lab <- deparse(substitute(x))
  ggplot() + 
  geom_point(data = df2_notime, aes(x = dist, y = x, color = pop),size=3) + 
  geom_line(data = df4_predict, aes(x = new_dist, y = y, linetype = as.factor(new_pop)),linewidth=1) + 
  theme_bw() + 
  labs(title = y.lab)}

```

```{r, include = F}

urlfile <- "https://raw.githubusercontent.com/kj-kling/Kling-etal-People-Nature/main/plots.csv"

df1 <- read_csv(url(urlfile)) # Download full botanical plot dataset from github. Both population & distance have been standardized (log(x + 1))


# Make Site a factor
df1$site <- factor(df1$site)

# Create dataset
# Includes botanical plots inside protected area & near all 13 communities sampled in 2015
df2_notime <- df1[df1$time==0,] # Data to use for analyses that do NOT involve plots repeat-sampled over time
df2_notime <- as.data.frame(df2_notime)

# Create dataset
# Includes botanical plots repeat-sampled between 2015-2021 near 5 communities
temp <- df1[df1$time>0,]
df3_time <- df1[df1$plot.id %in% temp$plot.id,] # Data to use for analyses of change over time
df3_time <- as.data.frame(df3_time)
```
######
## Modeling Population & Distance
##### *These analyses include botanical plots inside protected area & near all 13 communities sampled in 2015.*
######
### Timber & Medicinal Plant Analyses
#### TIMBER COUNT

```{r}

fun.mod.notime.aicc(8) # Use function to calculate aicc values of all possible models
aicc.w # Best model is Full-Int (w = 0.71)

mod.timc <-  glmmTMB(wood.ct ~ pop*dist + (1|site:plot.id), family = poisson, data = df2_notime)

summary(mod.timc)
```
######
##### How much more likely is this model than the null?
##### *First number below is dAICc between model & null & second is the actual likelihood (previous number/2 & exponentiated).*
```{r, echo = F}
fun.mod.e2.null(5)
```
######
##### Main effects: Timber Count
##### *Are the dist and population slopes significant if averaged over our nested random effects?*

```{r}
main.timc <- avg_slopes(mod.timc)
main.timc 
```
######
##### Marginal effects: Timber Count
##### *Does distance have a significant effect at certain population values?*
###### *Note: Population is standardized. Categories examined for marginal effects correspond to population values representative of our sample: (0 ppl, 50, 250, 2000)*

```{r}
me.timc <- ggpredict(mod.timc, terms = c("dist", "pop [0, 3.9, 5.5, 7.6]"), type = "fixed", allow.new.levels = T)
ggeffects::hypothesis_test(me.timc, test = NULL) 
```
######
##### *Note: All visualizations available either in main text of manuscript or in supplementary materials, as composite figures. These are just quick ones for your reference, as you go through the code!*
##### Visualize Data & Model-Predicted Results for Timber Count
###### *Note: Distance is standardized. Range corresponds to closest botanical plot distance to a community (~ 500 m) to ~ 22 km.*

```{r, echo = F}
# In order to visualize
# Creating data frame for predicted data from models for population & distance effects
# Data frame has distance levels from 6 (natural log of ~ 400 m) to 10 (~ 22000 m) & the following population levels (when exponentiated): 0 ppl (i.e., protected area), 50, 250, 2000
df.predict <- NULL
df.predict$dist <- rep(seq(from = 6,to = 10,by = 0.2),4*length(unique(df2_notime$site))*length(unique(df2_notime$plot.id)))
df.predict$pop <- rep(c(0, 3.9, 5.5, 7.6),each =
                         length(unique(df.predict$dist))*length(unique(df2_notime$site))*length(unique(df2_notime$plot.id)))
df.predict$site <- rep(unique(df2_notime$site),each = length(unique(df.predict$dist))*length(unique(df2_notime$plot.id)))
df.predict$plot.id <- rep(unique(df2_notime$plot.id),each = length(unique(df.predict$dist))*length(unique(df2_notime$site)))
df.predict <- as.data.frame(df.predict)

# FIRST, create data frame to house all predictions across models. Starting with that for timber count
df.predict$predict.timc <- exp(predict(mod.timc, newdata = df.predict, allow.new.levels = T))
df.predict$ave_predicted <- ave(df.predict$predict.timc, df.predict$dist, df.predict$pop, FUN=mean)
  
df4_predict <- unique(df.predict[,c("dist","pop","ave_predicted")])
colnames(df4_predict) <- c("new_dist","new_pop","tim.c")

fun.plot.mod(df2_notime$wood.ct, df4_predict$tim.c)
```

```{r}
```
######
#### TIMBER SPECIES RICHNESS

```{r}
## For remaining analyses of pop & dist, using the same process as that for TIMBER COUNT. Refer to Timber Count annotations

fun.mod.notime.aicc(11)
aicc.w # Best model is Full-Int (w = 0.74)

mod.timr <-  glmmTMB(wood.rich ~ pop*dist + (1|site:plot.id), family = poisson, data = df2_notime)

summary(mod.timr)
```
######
##### How much more likely is this model than the null?
```{r, echo = F}
fun.mod.e2.null(5)
```
######
##### Main effects: Timber Species Richness

```{r}
main.timr <- avg_slopes(mod.timr)
main.timr
```
######
##### Marginal effects: Timber Species Richness

```{r}
me.timr <- ggpredict(mod.timr, terms = c("dist", "pop [0, 3.9, 5.5, 7.6]"), type = "fixed", allow.new.levels = T) 
hypothesis_test(me.timr, test = NULL) 
```
######
##### Visualize Data & Model-Predicted Results for Timber Species Richness

```{r, echo = F}
df4_predict$tim.r <- fun.mod.predict(mod.timr)

fun.plot.mod(df2_notime$wood.rich, df4_predict$tim.r)
```

```{r}
```
######
#### MEDICINAL PLANTS COUNT

```{r}
fun.mod.notime.aicc(9)
aicc.w # Best model is Full-NoInt (w = 0.71)

mod.medc <- glmmTMB(med.ct ~ pop + dist + (1|site:plot.id), family = poisson, data = df2_notime)

summary(mod.medc) 
```
######
##### How much more likely is this model than the null?
```{r, echo = F}
fun.mod.e2.null(5)
```
######
##### Main effects: Medicinal Plants Count

```{r}
main.medc <- avg_slopes(mod.medc)
main.medc
```
######
##### Marginal effects: Medicinal Plants Count

```{r}
me.medc <- ggpredict(mod.medc, terms = c("dist", "pop [0, 3.9, 5.5, 7.6]"), type = "fixed", allow.new.levels = T) 
hypothesis_test(me.medc, test = NULL) 
```
######
##### Visualize Data & Model-Predicted Results for Medicinal Plants Count
###### *Note: There is no interaction between pop & dist for medicinal plants count.*

```{r, echo = F}
# REMEMBER, no interaction in the model
df4_predict$med.c <- fun.mod.predict(mod.medc)

fun.plot.mod(df2_notime$med.ct, df4_predict$med.c)
```

```{r}
```
######
#### MEDICINAL PLANTS SPECIES RICHNESS

```{r}
fun.mod.notime.aicc(12)
aicc.w # Best model is Pop only (w = 0.53)

mod.medr <- glmmTMB(med.rich ~ pop + (1|site:plot.id), family = poisson, data = df2_notime)

summary(mod.medr) 
```
######
##### How much more likely is this model than the null?
```{r, echo = F}
fun.mod.e2.null(5)
```
######
##### Main effects: Medicinal Plants Species Richness

```{r}
main.medr <- avg_slopes(mod.medr)
main.medr 
```
######
##### Marginal effects: Medicinal Plants Species Richness

```{r}
me.medr <- ggpredict(mod.medr, terms = c("pop [0, 3.9, 5.5, 7.6]"), type = "fixed", allow.new.levels = T) 
hypothesis_test(me.medr, test = NULL) 
```
######
##### Visualize Data & Model-Predicted Results for Medicinal Plants Species Richness
###### *Note: There is no distance term in the model for medicinal plants species richness.*

```{r, echo = F}
df4_predict$med.r <- fun.mod.predict(mod.medr)

fun.plot.mod(df2_notime$med.rich, df4_predict$med.r)
```

```{r}
```
######
######
### Lemur Food Trees Analyses
#### LEMUR FOOD TREES COUNT

```{r}
fun.mod.notime.aicc(10)
aicc.w # Best model is Full-Int (w = 0.64)

mod.lemc <- glmmTMB(lem.ct ~ pop*dist + (1|site:plot.id), family = poisson, data = df2_notime)

summary(mod.lemc)
```
######
##### How much more likely is this model than the null?
```{r, echo = F}
fun.mod.e2.null(5)
```

######
##### Main effects: Lemur Food Trees Count

```{r}
main.lemc <- avg_slopes(mod.lemc)
main.lemc 
```
######
##### Marginal effects: Lemur Food Trees Count

```{r}
me.lemc <- ggpredict(mod.lemc, terms = c("dist", "pop [0, 3.9, 5.5, 7.6]"), type = "fixed", allow.new.levels = T) 
hypothesis_test(me.lemc, test = NULL) 
```
######
##### Visualize Data & Model-Predicted Results for Lemur Food Trees Count

```{r, echo = F}
# Lemur Food Tree COUNT
df4_predict$lem.c <- fun.mod.predict(mod.lemc)

fun.plot.mod(df2_notime$lem.ct, df4_predict$lem.c)
```

```{r}
```
######
#### LEMUR FOOD TREES SPECIES RICHNESS

```{r}
# Lemur Food Tree SPECIES RICHNESS
fun.mod.notime.aicc(13)
aicc.w # Best model is Full-Int (w = 0.70)

mod.lemr <- glmmTMB(lem.rich ~ pop*dist + (1|site:plot.id), family = poisson, data = df2_notime)

summary(mod.lemr) 
```
######
##### How much more likely is this model than the null?
```{r, echo = F}
fun.mod.e2.null(5)
```

######
##### Main effects: Lemur Food Trees Species Richness

```{r}
main.lemr <- avg_slopes(mod.lemr)
main.lemr 
```
######
##### Marginal effects: Lemur Food Trees Species Richness

```{r}
me.lemr <- ggpredict(mod.lemr, terms = c("dist", "pop [0, 3.9, 5.5, 7.6]"), type = "fixed", allow.new.levels = T) 
hypothesis_test(me.lemr, test = NULL) 
```
######
##### Visualize Data & Model-Predicted Results for Lemur Food Trees Species Richness

```{r, echo = F}
df4_predict$lem.r <- fun.mod.predict(mod.lemr)

fun.plot.mod(df2_notime$lem.rich, df4_predict$lem.r)
```

```{r}
```
######
######
## Modeling Changes in Plot Metrics Over Time
##### *These analyses include all plots that were repeat-sampled near 5 communities between 2015-2021. Each plot was sampled 5x.*
###### *Note: All changes over time analyses use the same fixed effect model structure for population & distance as that used in the first set of analyses for that plant trait (e.g., TIMBER COUNT's model of population & distance included the following as its fixed effects: pop x dist.*
######
######
### Timber & Medicinal Plant Analyses
#### TIMBER COUNT

```{r}
mod.time.timc <- glmmTMB(wood.ct ~ pop*dist + time + (1+time|site:plot.id), family = poisson, data = df3_time) 

summary(mod.time.timc) 
```
######
##### How much more likely is this model than the null? What is its model weight?
```{r, echo = F}
fun.mod.e2.null.t(mod.time.timc, 8)
 
```

######
##### Main effects: Timber Count
##### *Is the effect of time (& dist and population for this subset of repeat-sampled plots) significant if averaged over our nested random effects?*

```{r}
main.time.timc <- avg_slopes(mod.time.timc)
main.time.timc
```
######
##### Visualize Data & Model-Predicted Results for Timber Count Change over Time BY SITE
###### *Note: Lines represent change over time for each individual plot per site.*

```{r, echo = F}
df3_time$timc <-exp(predict(mod.time.timc)) # Add predicted data to time data frame

ggplot(data = df3_time) + 
  geom_point(data = df3_time, aes(x = time, y = wood.ct, color = dist, group = plot.id),size=3) +
  geom_line(data = df3_time, aes(x = time, y = timc, group = plot.id)) + 
  facet_wrap(~site)
```

```{r}
```
######
#### TIMBER SPECIES RICHNESS

```{r}
mod.time.timr <- glmmTMB(wood.rich ~ pop*dist + time + (1+time|site:plot.id), family = poisson, data = df3_time) 

summary(mod.time.timr) 
```
##### How much more likely is this model than the null? What is its model weight?
```{r, echo = F}
fun.mod.e2.null.t(mod.time.timr, 11)
 
```

######
##### Main effects: Timber Species Richness

```{r}
main.time.timr <- avg_slopes(mod.time.timr)
main.time.timr
```
######
##### Visualize Data & Model-Predicted Results for Timber Species Richness Change over Time BY SITE

```{r, echo = F}
df3_time$timr <- exp(predict(mod.time.timr))

ggplot(data = df3_time) + 
  geom_point(data = df3_time, aes(x = time, y = wood.rich, color = dist, group = plot.id),size=3) +
  geom_line(data = df3_time, aes(x = time, y = timr, group = plot.id)) + 
  facet_wrap(~site)
```

```{r}
```
######
#### MEDICINAL PLANTS COUNT

```{r}
mod.time.medc <- glmmTMB(med.ct ~ pop + dist + time + (1+time|site:plot.id), family = poisson, data = df3_time)

summary(mod.time.medc) 
```
##### How much more likely is this model than the null? What is its model weight?
```{r, echo = F}
fun.mod.e2.null.t(mod.time.medc, 9)
 
```
######
##### Main effects: Medicinal Plants Count

```{r}
main.time.medc <- avg_slopes(mod.time.medc)
main.time.medc
```
######
##### Visualize Data & Model-Predicted Results for Medicinal Plants Count Change over Time BY SITE

```{r, echo = F}
df3_time$medc <- exp(predict(mod.time.medc))

ggplot(data = df3_time) + 
  geom_point(data = df3_time, aes(x = time, y = med.ct, color = dist, group = plot.id),size=3) +
  geom_line(data = df3_time, aes(x = time, y = medc, group = plot.id)) + 
  facet_wrap(~site)
```

```{r}
```
######
#### MEDICINAL PLANTS SPECIES RICHNESS
###### *Note: There is a singular convergence issue is time is included as a random slope. This is the only time analysis that does NOT include time as a random slope. Time is still included as a fixed effect.*

```{r}
mod.time.medr <- glmmTMB(med.rich ~ pop + time + (1|site:plot.id), family = poisson, data = df3_time) # Singular convergence issue when time was included as a random slope

summary(mod.time.medr) # Medicinal plant species richness significantly decreases over time
```
##### How much more likely is this model than the null? What is its model weight?
```{r, echo = F}
fun.mod.e2.null.t(mod.time.medr, 12)
 
```
######
##### Main effects: Medicinal Plants Species Richness

```{r}
main.time.medr <- avg_slopes(mod.time.medr)
main.time.medr
```
######
##### Visualize Data & Model-Predicted Results for Medicinal Plants Species Richness Change over Time BY SITE

```{r, echo = F}
df3_time$medr <- exp(predict(mod.time.medr))

ggplot(data = df3_time) + 
  geom_point(data = df3_time, aes(x = time, y = med.rich, color = dist, group = plot.id),size=3) +
  geom_line(data = df3_time, aes(x = time, y = medr, group = plot.id)) + 
  facet_wrap(~site)
```


```{r}
```
######
######
### Lemur Food Trees Analyses
#### LEMUR FOOD TREES COUNT

```{r}
mod.time.lemc <- glmmTMB(lem.ct ~ pop*dist + time + (1+time|site:plot.id), family = poisson, data = df3_time) 

summary(mod.time.lemc) 
```
##### How much more likely is this model than the null? What is its model weight?
```{r, echo = F}
fun.mod.e2.null.t(mod.time.lemc, 10)
 
```
######
##### Main effects: Medicinal Plants Species Richness

```{r}
main.time.lemc <- avg_slopes(mod.time.lemc)
main.time.lemc
```
######
##### Visualize Data & Model-Predicted Results for Lemur Food Trees Count Change over Time BY SITE

```{r, echo = F}
df3_time$lemc <- exp(predict(mod.time.lemc))

ggplot(data = df3_time) + 
  geom_point(data = df3_time, aes(x = time, y = lem.ct, color = dist, group = plot.id),size=3) +
  geom_line(data = df3_time, aes(x = time, y = lemc, group = plot.id)) + 
  facet_wrap(~site)
```

```{r}
```
######
#### LEMUR FOOD TREES SPECIES RICHNESS

```{r}
mod.time.lemr <- glmmTMB(lem.rich ~ pop*dist + time + (1+time|site:plot.id), family = poisson, data = df3_time) 

summary(mod.time.lemr)
```
##### How much more likely is this model than the null? What is its model weight?
```{r, echo = F}
fun.mod.e2.null.t(mod.time.lemr, 13)
 
```
######
##### Main effects: Lemur Food Trees Species Richness

```{r}
main.time.lemr <- avg_slopes(mod.time.lemr)
main.time.lemr
```
######
##### Visualize Data & Model-Predicted Results for Lemur Food Trees Species RIchness Change over Time BY SITE

```{r, echo = F}
df3_time$lemr <- exp(predict(mod.time.lemr))

ggplot(data = df3_time) + 
  geom_point(data = df3_time, aes(x = time, y = lem.rich, color = dist, group = plot.id),size=3) +
  geom_line(data = df3_time, aes(x = time, y = lemr, group = plot.id)) + 
  facet_wrap(~site)
```

```{r}
```

######
## Bonus Plot!
##### *Plot showing Incidence Rate Ratios for all models. Note that plant metrics were not standardized and thus are not directly comparable.*

```{r, echo = F}

est <- read.csv("est.plottab.csv", header = T)
est$type <- as.factor(est$type)
est$type <- factor(est$type, levels = c("Timber", "Medicinal Plant", "Ruffed Lemur Food Tree"))

# Estimate (IRR) Plot
forestplot(df = est, name = var, estimate = estimate, pvalue = p.val, psignif = 0.05, logodds = T,
           colour = type, fatten = 2,
           xlab = "\nIncidence Rate Ratios\nfor Plot Count & Species Richness (95% CI)") +
  ggforce::facet_col(facets = ~ var2, scales = "free_y", space = "free") +
  theme(#plot.title = element_text(hjust = 0.5),
    #plot.subtitle = element_text(hjust = 0.5),
    #axis.text.y = element_text(size = rel(1.2)),
    strip.background = element_rect(color = "black", fill = NA, size = 1.25),
    legend.title = element_blank(),
    legend.text = element_text(size = rel(1)),
    legend.position = "top",
    legend.justification="right",
    legend.margin=margin(0,0,0,0),
    legend.box.margin=margin(-5,-5,-5,-5)) +
  guides(color = guide_legend(override.aes = list(size = 1))) + 
  geom_effect(aes(xmin = .data$.xmin, xmax = .data$.xmax, colour = type, filled = .data$.filled), 
              fatten = 3.5) +
  scale_color_manual(values = c("#44aa99", "#ddcc77", "#aa4499")) 